---
title: "Exploratory Graph Analysis of the DBAS-16"
subtitle: "Example Rmd script"
author: "Author: [name removed for blinded review]"
date: " `r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    tables: true
bibliography: ["../ext/bib/references.bib", "../ext/bib/r-references.bib"]
csl: ../ext/apa7.csl
---

```{r packages-nwk, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(psychTools)
library(EGAnet)
library(qgraph)
library(tidymodels)
library(ggpubr)
```

```{r dataset-ega, warning=FALSE, message=FALSE, echo=FALSE}
## read full dataset
mydata <- read.csv("../data/data.csv")

## select answers from arm 1
dbas <- mydata |> 
  dplyr::filter(redcap_event_name == "elegibilidade_arm_1",
                ## exclude previously identified outliers 
                !record_id %in% c(1651, 2015, 2938, 3793)) |>
  dplyr::mutate(group = dplyr::case_when(
    dsm_1 == 0 &
    dsm_2 == 0 &
    dsm_3 == 0 &
    dsm_4 == 0 &
    dsm_5 == 0 &
    isi_total < 8 ~ "good_sleepers",
    TRUE ~ "bad_sleepers")) |> 
  dplyr::select(record_id, paste0("dbas16_", 1:16), group) |> 
  dplyr::filter(!dplyr::if_all(dplyr::starts_with("dbas16_"), ~ is.na(.))) |> 
  # median inputation for one participant that left one item unanswered
  dplyr::mutate(dbas16_10 = tidyr::replace_na(dbas16_10, 7))

items <- dbas[, 2:17]

colnames(items) <- paste0("D", 1:16)
```

```{r sample-split}
set.seed(1354789)

dbas_initial_split <- dbas |>  initial_split(1/2, strata=group)

dbas_train <- training(dbas_initial_split)[,2:17]
dbas_test <- testing(dbas_initial_split)[,2:17]

colnames(dbas_train) <- colnames(dbas_test) <- paste0("D", 1:16)
```



```{r labels}
items_labels <- tibble::tribble(
  ~items, ~description,
  "D1" , "I need 8 hours of sleep to feel refreshed and function well during the day.",
  "D2" , "When I don't get proper amount of sleep on a given night, I need to catch up on the next day by napping or on the next night by sleeping longer.",
  "D3" ,"I am concerned that chronic insomnia may have serious consequences on my physical health.",
  "D4" ,"I am worried that I may lose control over my abilities to sleep.",
  "D5" ,"After a poor night’s sleep, I know that it will interfere with my daily activities on the next day.",
  "D6" ,"In order to be alert and function well during the day, I believe I would be better off taking a sleeping pill rather than having a poor night’s sleep.",
  "D7" ,"When I feel irritable, depressed, or anxious during the day, it is mostly because I did not sleep well the night before.",
  "D8" ,"When I sleep poorly on one night, I know it will disturb my sleep schedule for the whole week.",
  "D9" ,"Without an adequate night’s sleep, I can hardly function the next day.",
  "D10" ,"I can’t ever predict whether I’ll have a good or poor night’s sleep.",
  "D11" ,"I have little ability to manage the negative consequences of disturbed sleep.",
  "D12" ,"When I feel tired, have no energy, or just seem not to function well during the day, it is generally because I did not sleep well the night before.",
  "D13" ,"I believe insomnia is essentially the result of a chemical imbalance.",
  "D14" ,"I feel insomnia is ruining my ability to enjoy life and prevents me from doing what I want.",
  "D15" ,"Medication is probably the only solution to sleeplessness.",
  "D16" ,"I avoid or cancel obligations (social, family) after a poor night’s sleep."
  
)
```


# Derivation

# Detecting redundancy with Unique Variable Analysis

```{r redund-items}
redund <- EGAnet::UVA(data=dbas_train, method="wTO", type="threshold",  corr =  "spearman",
                    reduce=FALSE, adhoc=FALSE, plot.redundancy=TRUE) 
redund.items <- redund$redundancy$redundant
```

Redundancies were found between items `r names(redund.items[1])` and `r redund.items[[1]]`, and between items `r names(redund.items[2])` and `r redund.items[[2]]`.

```{r}
psych::describe(dbas_train[, c("D1", "D2","D3", "D4", "D6", "D15")]) |> 
  as.data.frame() |> 
  dplyr::mutate(maxwTO = c(
    max(redund$redundancy$weights[1,-2]),
    max(redund$redundancy$weights[2,-1]),
    max(redund$redundancy$weights[3,-4]),
    max(redund$redundancy$weights[4,-3]),
    max(redund$redundancy$weights[6,-15]),
    max(redund$redundancy$weights[15,-6])
  ),
  item_total_cor = psych::alpha(dbas_train)$item.stats[c(1, 2, 3, 4, 6, 15), 4])
```


# Initial Dimensionality Estimation

## Model with redundancies

```{r ega-with-redundancy}
# EGA (with redundancy) 
ega.wr.tmfg <- EGA(dbas_train, algorithm = "walktrap", model = "TMFG",  corr =  "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

ega.wr.glasso <- EGA(dbas_train, algorithm = "walktrap", model = "glasso", corr =  "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels, plot.args = list()) 


comp.plots.wr <- EGAnet::compare.EGA.plots(
  ega.wr.glasso, ega.wr.tmfg,
  base.plot = 1,
  labels = c("GLASSO", "TMFG"),
  rows = 1, columns = 2
)

comp.plots.wr$comparison.plot

```


```{r dim-tab-wr}
dimtab.wr <- ega.wr.tmfg$dim.variables |> 
  dplyr::left_join(ega.wr.glasso$dim.variables, by = "items", suffix = c("_tmfg", "_glasso"))
dimtab.wr
```
### Total Entropy Fit Index


```{r tefi}

dbas.cor <- abs(qgraph::cor_auto(dbas_train))

tefi_tmfg <- tefi(data=dbas.cor, structure=ega.wr.tmfg$wc) 

tefi_glasso <- tefi(data=dbas.cor, structure=ega.wr.glasso$wc) 

cbind("Algorithm" = c("TMFG", "GLASSO"), rbind(tefi_tmfg, tefi_glasso))

```

TEFI for the TFMG model is `r tefi_tmfg$VN.Entropy.Fit`, and for the GLASSO model is `r tefi_glasso$VN.Entropy.Fit`, making the model with 4 dimensions preferable.

### Structural consistency


```{r bootEGA-wr}
# Uncomment to run the functions

# Compute bootstrap
# boot.tmfg.wr <- bootEGA(dbas_train, iter = 1000, model = "TMFG", type = "parametric", corr =  "spearman", uni.method = "LE", plot.typicalStructure = FALSE)
# 
# boot.glasso.wr <- bootEGA(dbas_train, iter = 1000, model = "glasso", type = "parametric", corr =  "spearman", uni.method = "LE", plot.typicalStructure = FALSE)
# 
# saveRDS(boot.tmfg.wr, "../output/ega/bootTMFGwr.rds")
# saveRDS(boot.glasso.wr, "../output/ega/bootGLASSOwr.rds")

boot.tmfg.wr <-readRDS("../output/ega/bootTMFGwr.rds")
boot.glasso.wr <-readRDS("../output/ega/bootGLASSOwr.rds")
```

```{r stability-descriptive}
# Descriptive statistics 
desc.stab.wr <- rbind(boot.tmfg.wr$summary.table, boot.glasso.wr$summary.table)
row.names(desc.stab.wr) <- c("TMFG", "GLASSO")
desc.stab.wr
```

```{r dimension-frequency}
# Frequency of dimensions 

dim.freq.wr <- rbind(boot.tmfg.wr$frequency, boot.glasso.wr$frequency)
row.names(dim.freq.wr) <- c(rep("TMFG", nrow(boot.tmfg.wr$frequency)), 
                            rep("GLASSO", nrow(boot.glasso.wr$frequency)))
dim.freq.wr
```


```{r structural-consistency-wr}
# structural consistency
dim.stab.wr.tmfg <- dimensionStability(boot.tmfg.wr)
dim.stab.wr.glasso <- dimensionStability(boot.glasso.wr)
```

```{r dim-stab-wr}
dim.stab.wr <- tibble::enframe(list(
  "tmfg" = dim.stab.wr.tmfg$dimension.stability$structural.consistency,
"glasso" = dim.stab.wr.glasso$dimension.stability$structural.consistency)) |> 
  tidyr::unnest(value) |> 
  dplyr::with_groups(name, mutate, dim = row_number(), .before="value")
dim.stab.wr
```

### Standardized node strenght

```{r network-loadings-wr}
# Compute standardized node strength 
net.loads(ega.wr.glasso)$std
```


```{r}
# Item stability across all dimensions
dim.stab.wr.tmfg$item.stability$item.stability$all.dimensions
```

## EGA without item D15

```{r ega-d15}
# EGA without D15
ega.d15.tmfg <- EGA(dbas_train[,-c(15)], algorithm = "walktrap", model = "TMFG", corr = "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

ega.d15.glasso <- EGA(dbas_train[,-c(15)], algorithm = "walktrap", model = "glasso", corr = "spearman", 
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

comp.plots.d15 <- EGAnet::compare.EGA.plots(
  ega.d15.glasso, ega.d15.tmfg,
  base.plot = 1,
  labels = c("GLASSO", "TMFG"),
  rows = 1, columns = 2
)

comp.plots.d15$comparison.plot
```


```{r dim-tab-d15}
dimtab.d15 <- ega.d15.tmfg$dim.variables |> 
  dplyr::left_join(ega.d15.glasso$dim.variables, by = "items", suffix = c("_tmfg", "_glasso"))
dimtab.d15
```

### Total Entropy Fit Index


```{r tefi-d15}

dbas.cor.d15 <- abs(qgraph::cor_auto(dbas_train[,-c(15)]))

tefi_tmfg.d15 <- tefi(data=dbas.cor.d15, structure=ega.d15.tmfg$wc) 

tefi_glasso.d15 <- tefi(data=dbas.cor.d15, structure=ega.d15.glasso$wc) 

cbind("Algorithm" = c("TMFG", "GLASSO"), rbind(tefi_tmfg.d15, tefi_glasso.d15))

```
### Structural consistency

<!-- un-comment to recompute the bootstrap analyses. results were saved only for rendering the manuscript faster -->

```{r bootEGA-d15}
# Compute bootstrap 
# boot.tmfg.d15 <- bootEGA(dbas_train[,-c(15)], iter = 1000, model = "TMFG", type = "parametric", corr = "spearman", plot.typicalStructure = FALSE, uni.method = "LE")
# 
# boot.glasso.d15 <- bootEGA(dbas_train[,-c(15)], iter = 1000, model = "glasso", type = "parametric", corr = "spearman", plot.typicalStructure = FALSE, uni.method = "LE")
# 
# saveRDS(boot.tmfg.d15, "../output/ega/bootTMFGd15.rds")
# saveRDS(boot.glasso.d15, "../output/ega/bootGLASSOd15.rds")

boot.tmfg.d15 <-readRDS("../output/ega/bootTMFGd15.rds")
boot.glasso.d15 <-readRDS("../output/ega/bootGLASSOd15.rds")
```

```{r stability-descriptive-d15}
# Descriptive statistics 
desc.stab.d15 <- rbind(boot.tmfg.d15$summary.table, boot.glasso.d15$summary.table)
row.names(desc.stab.d15) <- c("TMFG", "GLASSO")
desc.stab.d15
```

```{r dimension-frequency-d15}
# Frequency of dimensions 

dim.freq.d15 <- rbind(boot.tmfg.d15$frequency, boot.glasso.d15$frequency)
row.names(dim.freq.d15) <- c(rep("TMFG", nrow(boot.tmfg.d15$frequency)), 
                            rep("GLASSO", nrow(boot.glasso.d15$frequency)))
dim.freq.d15
```


```{r structural-consistency-d15}
# structural consistency
dim.stab.d15.tmfg <- dimensionStability(boot.tmfg.d15)
dim.stab.d15.glasso <- dimensionStability(boot.glasso.d15)
```

```{r dim-stab-d15}
dim.stab.d15 <- tibble::enframe(list(
  "tmfg" = dim.stab.d15.tmfg$dimension.stability$structural.consistency,
"glasso" = dim.stab.d15.glasso$dimension.stability$structural.consistency)) |> 
  tidyr::unnest(value) |> 
  dplyr::with_groups(name, mutate, dim = row_number(), .before="value")

```

```{r}
# Item stability across all dimensions
dim.stab.d15.glasso$item.stability$item.stability$all.dimensions
```

## EGA with items D15 and D3 removed

```{r ega-d3}
# EGA without D3
ega.d3.tmfg <- EGA(dbas_train[,-c(15,3)], algorithm = "walktrap", model = "TMFG", corr = "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

ega.d3.glasso <- EGA(dbas_train[,-c(15,3)], algorithm = "walktrap", model = "glasso", corr = "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

comp.plots.d3 <- EGAnet::compare.EGA.plots(
  ega.d3.glasso, ega.d3.tmfg,
  base.plot = 1,
  labels = c("GLASSO", "TMFG"),
  rows = 1, columns = 2
)

comp.plots.d3$comparison.plot
```


```{r dim-tab-d3}
dimtab.d3 <- ega.d3.tmfg$dim.variables |> 
  dplyr::left_join(ega.d3.glasso$dim.variables, by = "items", suffix = c("_tmfg", "_glasso"))
dimtab.d3
```

### Total Entropy Fit Index


```{r tefi-wo-d3}

dbasD3.cor <- abs(qgraph::cor_auto(dbas_train[,-c(15,3)]))

tefi_tmfgD3 <- tefi(data=dbasD3.cor, structure=ega.d3.tmfg$wc) 

tefi_glassoD3 <- tefi(data=dbasD3.cor, structure=ega.d3.glasso$wc) 

cbind("Algorithm" = c("TMFG", "GLASSO"), rbind(tefi_tmfgD3, tefi_glassoD3))

```


### Structural consistency


```{r bootEGA-d3}

# boot.tmfg.d3 <- bootEGA(dbas_train[,-c(15,3)], iter = 1000, model = "TMFG", type = "parametric", corr = "spearman", uni.method = "LE", plot.typicalStructure = FALSE)
# 
# boot.glasso.d3 <- bootEGA(dbas_train[,-c(15,3)], iter = 1000, model = "glasso", type = "parametric", corr = "spearman", uni.method = "LE", plot.typicalStructure = FALSE)
# 
# saveRDS(boot.tmfg.d3, "../output/ega/bootTMFGd3.rds")
# saveRDS(boot.glasso.d3, "../output/ega/bootGLASSOd3.rds")

boot.tmfg.d3 <-readRDS("../output/ega/bootTMFGd3.rds")
boot.glasso.d3 <-readRDS("../output/ega/bootGLASSOd3.rds")
```

```{r stability-descriptive-d3}
# Descriptive statistics 
desc.stab.d3 <- rbind(boot.tmfg.d3$summary.table, boot.glasso.d3$summary.table)
row.names(desc.stab.d3) <- c("TMFG", "GLASSO")
desc.stab.d3
```

```{r dimension-frequency-d3}
# Frequency of dimensions 

dim.freq.d3 <- rbind(boot.tmfg.d3$frequency, boot.glasso.d3$frequency)
row.names(dim.freq.d3) <- c(rep("TMFG", nrow(boot.tmfg.d3$frequency)), 
                            rep("GLASSO", nrow(boot.glasso.d3$frequency)))
dim.freq.d3
```

```{r structural-consistency-d3}
# structural consistency
dim.stab.d3.tmfg <- dimensionStability(boot.tmfg.d3)
dim.stab.d3.glasso <- dimensionStability(boot.glasso.d3)
```

```{r dim-stab-d3}
dim.stab.d3 <- tibble::enframe(list(
  "tmfg" = dim.stab.d3.tmfg$dimension.stability$structural.consistency,
"glasso" = dim.stab.d3.glasso$dimension.stability$structural.consistency)) |> 
  tidyr::unnest(value) |> 
  dplyr::with_groups(name, mutate, dim = row_number(), .before="value")
dim.stab.d3
```

## EGA with items D15, D3, and D1 removed

```{r ega-d1-tmfg}

ega.d1.tmfg <- EGA(dbas_train[,-c(15, 3, 1)], algorithm = "walktrap", model = "TMFG", corr = "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

ega.d1.glasso <- EGA(dbas_train[,-c(15, 3, 1)], algorithm = "walktrap", model = "glasso", corr = "spearman",
              plot.EGA = FALSE, uni.method = "LE", key = items_labels) 

comp.plots.d1 <- EGAnet::compare.EGA.plots(
  ega.d1.glasso, ega.d1.tmfg,
  base.plot = 1,
  labels = c("GLASSO", "TMFG"),
  rows = 1, columns = 2
)

comp.plots.d1$comparison.plot
```

```{r dim-tab-d1}
dimtab.d1 <- ega.d1.tmfg$dim.variables |> 
  dplyr::left_join(ega.d1.glasso$dim.variables, by = "items", suffix = c("_tmfg", "_glasso"))
dimtab.d1
```

### Total Entropy Fit Index


```{r tefi-wo-d1}

dbas.d1.cor <- abs(qgraph::cor_auto(dbas_train[,-c(15, 3, 1)]))

tefi.tmfg.d1 <- tefi(data=dbas.d1.cor, structure=ega.d1.tmfg$wc) 

tefi.glasso.d1 <- tefi(data=dbas.d1.cor, structure=ega.d1.glasso$wc) 

cbind("Algorithm" = c("TMFG", "GLASSO"), rbind(tefi.tmfg.d1, tefi.glasso.d1))

```


### Structural consistency

```{r bootEGA-d1}

# boot.tmfg.d1 <- bootEGA(dbas_train[,-c(15, 3, 1)], iter = 1000, model = "TMFG", type = "parametric", corr = "spearman", uni.method = "LE", plot.typicalStructure = FALSE, progress=FALSE)
# 
# boot.glasso.d1 <- bootEGA(dbas_train[,-c(15, 3, 1)], iter = 1000, model = "glasso", type = "parametric", corr = "spearman", uni.method = "LE", plot.typicalStructure = FALSE, progress=FALSE)
# 
# saveRDS(boot.tmfg.d1, "../output/ega/bootTMFGd1.rds")
# saveRDS(boot.glasso.d1, "../output/ega/bootGLASSOd1.rds")

boot.tmfg.d1 <-readRDS("../output/ega/bootTMFGd1.rds")
boot.glasso.d1 <-readRDS("../output/ega/bootGLASSOd1.rds")
```

```{r stability-descriptive-d1}
# Descriptive statistics 
desc.stab.d1 <- rbind(boot.tmfg.d1$summary.table, boot.glasso.d1$summary.table)
row.names(desc.stab.d1) <- c("TMFG", "GLASSO")
desc.stab.d1
```

```{r dimension-frequency-d1}
# Frequency of dimensions 

dim.freq.d1 <- rbind(boot.tmfg.d1$frequency, boot.glasso.d1$frequency)
row.names(dim.freq.d1) <- c(rep("TMFG", nrow(boot.tmfg.d1$frequency)), 
                            rep("GLASSO", nrow(boot.glasso.d1$frequency)))
dim.freq.d1
```

```{r structural-consistency-d1}
# structural consistency
dim.stab.d1.tmfg <- dimensionStability(boot.tmfg.d1)
dim.stab.d1.glasso <- dimensionStability(boot.glasso.d1)
```

```{r dim-stab-d1}
dim.stab.d1 <- tibble::enframe(list(
  "tmfg" = dim.stab.d1.tmfg$dimension.stability$structural.consistency,
"glasso" = dim.stab.d1.glasso$dimension.stability$structural.consistency)) |> 
  tidyr::unnest(value) |> 
  dplyr::with_groups(name, mutate, dim = row_number(), .before="value")
dim.stab.d1
```

## Model comparison

```{r compare}
cbind("Algorithm" = c("TMFG D15", "GLASSO D15", "TMFG D15-D3", "GLASSO D15-D3", "TMFG D15-D3-D1", "GLASSO D15-D3-D1","TMFG WR", "GLASSO WR"), rbind(tefi_tmfg.d15, tefi_glasso.d15, tefi_tmfgD3, tefi_glassoD3, tefi.tmfg.d1, tefi.glasso.d1, tefi_tmfg, tefi_glasso))
```

# Confirmation

## Redundancy

```{r redund-items-confirmation}
redund_test <- EGAnet::UVA(data=dbas_test, method="wTO", type="threshold",  corr =  "spearman",
                    reduce=FALSE, adhoc=FALSE, plot.redundancy=TRUE) 
redund.items.test <- redund_test$redundancy$redundant
```


## EGA with test sample

```{r ega-tmfg-test}

ega.test.tmfg <- EGA(dbas_test[,-c(15, 3, 1)], algorithm = "walktrap", model = "TMFG", corr = "spearman",
              plot.EGA = FALSE, key = items_labels) 

ega.test.glasso <- EGA(dbas_test[,-c(15, 3, 1)], algorithm = "walktrap", model = "glasso", corr = "spearman",
              plot.EGA = FALSE, key = items_labels) 

comp.plots.test <- EGAnet::compare.EGA.plots(
  ega.test.glasso, ega.test.tmfg,
  base.plot = 1,
  labels = c("GLASSO", "TMFG"),
  rows = 1, columns = 2
)

comp.plots.test$comparison.plot
```

```{r dim-tab-test}
dimtab.test <- ega.test.tmfg$dim.variables |> 
  dplyr::left_join(ega.test.glasso$dim.variables, by = "items", suffix = c("_tmfg", "_glasso"))
dimtab.test
```

### Total Entropy Fit Index


```{r tefi-test}

dbas.test.cor <- abs(qgraph::cor_auto(dbas_test[,-c(15, 3, 1)]))

tefi.tmfg.test <- tefi(data=dbas.test.cor, structure=ega.test.tmfg$wc) 

tefi.glasso.test <- tefi(data=dbas.test.cor, structure=ega.test.glasso$wc) 

cbind("Algorithm" = c("TMFG", "GLASSO"), rbind(tefi.tmfg.test, tefi.glasso.test))

```


### Structural consistency

```{r bootEGA-test}

# boot.tmfg.test <- bootEGA(dbas_test[,-c(15, 3, 1)], iter = 1000, model = "TMFG", type = "parametric", corr = "spearman", plot.typicalStructure = FALSE, progress=FALSE)
# 
# boot.glasso.test <- bootEGA(dbas_test[,-c(15, 3, 1)], iter = 1000, model = "glasso", type = "parametric", corr = "spearman", plot.typicalStructure = FALSE, progress=FALSE)
# 
# saveRDS(boot.tmfg.test, "../output/ega/bootTMFGtest.rds")
# saveRDS(boot.glasso.test, "../output/ega/bootGLASSOtest.rds")

boot.tmfg.test <-readRDS("../output/ega/bootTMFGtest.rds")
boot.glasso.test <-readRDS("../output/ega/bootGLASSOtest.rds")

```

```{r stability-descriptive-test}
# Descriptive statistics 
desc.stab.test <- rbind(boot.tmfg.test$summary.table, boot.glasso.test$summary.table)
row.names(desc.stab.test) <- c("TMFG", "GLASSO")
desc.stab.test
```

```{r dimension-frequency-test}
# Frequency of dimensions 

dim.freq.test <- rbind(boot.tmfg.test$frequency, boot.glasso.test$frequency)
row.names(dim.freq.test) <- c(rep("TMFG", nrow(boot.tmfg.test$frequency)), 
                            rep("GLASSO", nrow(boot.glasso.test$frequency)))
dim.freq.test
```

```{r structural-consistency-test}
# structural consistency
dim.stab.test.tmfg <- dimensionStability(boot.tmfg.test)
dim.stab.test.glasso <- dimensionStability(boot.glasso.test)
```

```{r dim-stab-test}
dim.stab.test <- tibble::enframe(list(
  "tmfg" = dim.stab.test.tmfg$dimension.stability$structural.consistency,
"glasso" = dim.stab.test.glasso$dimension.stability$structural.consistency)) |> 
  tidyr::unnest(value) |> 
  dplyr::with_groups(name, mutate, dim = row_number(), .before="value")
dim.stab.test
```
# EGA x CFA

```{r }

cfa.ega <- CFA(ega.obj = ega.d1.glasso, data = dbas_test, estimator="MLR", plot.CFA = FALSE)

ega.theory <- ega.wr.glasso

ega.theory$dim.variables[,1] <- colnames(items)

ega.theory$dim.variables[c(5, 7, 9, 12, 16),2] <- rep(1, 5)
ega.theory$dim.variables[c(3, 4, 8, 10, 11, 14),2] <- rep(2, 6)
ega.theory$dim.variables[c(1, 2),2] <- rep(3, 2)
ega.theory$dim.variables[c(6, 13, 15),2] <- rep(4, 3)

cfa.theory <- CFA(ega.obj = ega.theory, data = dbas_test, estimator="MLR", plot.CFA = FALSE)
```

```{r fit-comparison}
rbind("Theory model" = cfa.theory$fit.measures, 
      "EGA model" = cfa.ega$fit.measures
      )
```

```{r cfa-models}
mod_theory <- '
cons =~ D5 + D7 + D9 + D12 + D16
worry =~ D3 + D4 + D8 + D10 + D11 + D14
exp =~ D1 + D2
med =~ D6 + D13 + D15
'
mod_mi <- '
cons =~ D5 + D7 + D9 + D12 + D16
worry =~ D3 + D4 + D8 + D10 + D11 + D14
exp =~ D1 + D2
med =~ D6 + D13 + D15
D3 ~~  D4
D6 ~~ D15
'

mod_glasso <- '
F1 =~ D2 + D5 + D7 + D8 + D9 + D12 + D16
F2 =~ D4 + D6 + D10 + D11 + D13 + D14
'

  
fit_theory <- lavaan::cfa(model = mod_theory,
                        std.lv = TRUE, estimator = 'MLMV', data = dbas_test)

fit_mi <- lavaan::cfa(model = mod_mi,
                        std.lv = TRUE, estimator = 'MLMV', data = dbas_test)

fit_glasso <- lavaan::cfa(model = mod_glasso,
                        std.lv = TRUE, estimator = 'MLMV', data = dbas_test)

```

```{r LRT-models}
theory_glasso_comparison <- lavaan::lavTestLRT(fit_glasso, fit_theory, method="satorra.bentler.2010")
mi_glasso_comparison <- lavaan::lavTestLRT(fit_glasso, fit_mi, method="satorra.bentler.2010")
theory_glasso_comparison
```

```{r fit-measures-comparison}
fm_theory <- semTools::fitmeasures(fit_theory, 
                         fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi",
                                          "rmsea.robust", "rmsea.ci.lower.robust", "rmsea.ci.upper.robust", "tli.robust",
                                          "cfi.robust", "srmr_bentler"))

fm_mi <- semTools::fitmeasures(fit_mi, 
                         fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi",
                                          "rmsea.robust", "rmsea.ci.lower.robust", "rmsea.ci.upper.robust", "tli.robust",
                                          "cfi.robust", "srmr_bentler"))

fm_glasso <- semTools::fitmeasures(fit_glasso, 
                         fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi",
                                          "rmsea.robust", "rmsea.ci.lower.robust", "rmsea.ci.upper.robust", "tli.robust",
                                          "cfi.robust", "srmr_bentler"))


purrr::map_dfr(list("Theory" = fm_theory, "Theory w/ cor errors" = fm_mi, "EGA GLASSO" = fm_glasso), enframe, .id = "Model") |> 
  tidyr::pivot_wider(names_from = "Model", values_from = value)

```
```{r fit-differences}
theory_glasso <- fm_theory - fm_glasso
theory_glasso_diffs <- t(data.frame(theory_glasso))[, c("rmsea", "cfi", "rmsea.robust", "cfi.robust", "srmr_bentler")]
theory_glasso_diffs
```

## Internal consistency EGA model

```{r ega-lv-factors}
ega_factors <- list(
  dim1 = items[, c(2,5,7,8,9,12,16)],
  dim2 = items[, c(4,6,10,11,13,14)]
)
```

```{r ega-lv-reliability}
# library(furrr)
# future::plan(future::multisession)
# alpha_ega <- progressr::with_progress({
#   factors <- 1:10
#   p <- progressr::progressor(length(factors))
#   furrr::future_map(
#     ega_factors,
#     ~MBESS::ci.reliability(.x, type = "alpha", interval.type = "bca"),
#     prog = p,
#     .options = furrr_options(seed = 123)
#   )
# })
# 
# omega_ega <- purrr::map(ega_factors,
#          ~MBESS::ci.reliability(.x, type="hierarchical", interval.type = "perc", B = 1000))
# 
# saveRDS(alpha_ega, "../output/ega/alpha_ega.rds")
# saveRDS(omega_ega, "../output/ega/omega_ega.rds")

alpha_ega <- readRDS("../output/ega/alpha_ega.rds")
omega_ega <- readRDS("../output/ega/omega_ega.rds")
```



