---
main_topsize: 0.1 #percent coverage of the poster
main_bottomsize: 0.07
poster_height: "150cm"
poster_width: "80cm"
body_textsize: "52px"
#ESSENTIALS
title: '**Translation and validation of the Sleep Problem Acceptance Questionnaire (SPAQ) with a Brazilian sample**'
author:
  - name: '**Marwin M. I. B. Carmo, MSc**'
    affil: 3
    main: true
    #orcid: '0000-0002-1099-3857'
    email: mmcarmo@ucdavis.edu
  - name: Leo Paulos-Guarnieri
    affil: 1
    main: true
    email: 'guarnieri.leo@usp.br'
  - name: Maria Laura Nogueira Pires, PhD
    affil: 2
    main: true
    email: 'laurapires@actigrafia.org'
  - name: Renatha El Rafihi Ferreira, PhD
    affil: 1
    main: true
    email: 'rerafihi@usp.br'
affiliation:
  - num: 1
    address: Department of Psychiatry, University of São Paulo, Brazil
  - num: 2
    address: Private practice in sleep psychology
  - num: 3
    address: Department of Psychology, University of California, Davis
# Main Colour Scheme
primary_colour: "#024F84"
secondary_colour: "#C00000"
body_bgcol: "#ffffff"
#---POSTER BODY OPTIONS---#
reference_textsize: "38px"
caption_textsize: "38px"
affiliation_textsize: "38px"
#Column
column_numbers: 2
#Main Section
main_fontfamily: "PT Sans"
main_textalign: "left"
main_textsize: "160px"
main_textalpha: "cf"
main_findings:
  - "Preliminary evidence for a new measure \nof **Sleep Acceptance**: a promising tool for clinicians and researchers."
main_textcol: "#ffffff"
logoleft_name: img/fapesp-white.png
logoright_name: img/ipq-white3.png
logocenter_name: img/qrcodeSAS.svg
output: 
  posterdown::posterdown_betterport:
    highlight: NULL
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
    template: custom_template.html
#bibliography: references.bib
link-citations: true
#csl: apa7.csl
#nocite: '@*'
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
#options(knitr.table.format = "html") 
#options(kableExtra.html.bsTable = T)
```

```{r data, include=FALSE}
library(dplyr)
library(ggplot2)
library(modelsummary)
library(kableExtra)
library(psych)
library(lavaan)
library(MBESS)
library(tidyr)
library(semTools)
library("stringr")
library("semPlot")
library("corrr")
```

```{r dataset, include=FALSE}
## read full dataset
mydata <- read.csv("data.csv")
  

convdata <- mydata |> 
  dplyr::filter(!record_id %in% c(1651, 2015, 2938, 3793)) |>  # dupes
  dplyr::filter(redcap_event_name == "elegibilidade_arm_1",
    !dplyr::if_all(dplyr::starts_with("spaq_"), ~ is.na(.))) |> 
  dplyr::select(record_id, paste0("spaq_", 1:8), dplyr::starts_with("isi_"),
                dplyr::starts_with("aaq_"), dplyr::starts_with("dbas16_"),
                dplyr::starts_with("hads_")) |> 
  dplyr::mutate(across(spaq_5:spaq_8, ~ 6 - .x))

## select answers from arm 1
arm1 <- convdata |> 
  dplyr::select(record_id, paste0("spaq_", 1:8))

  
```


```{r, message=FALSE, warning=FALSE, include=FALSE}
# CFA ####

mod.spaq <- '
AE =~ spaq_1 + spaq_2 + spaq_3 + spaq_4 
WILL =~ spaq_5 + spaq_6 + spaq_7 + spaq_8 

'
fit.spaq <- lavaan::cfa(model = mod.spaq, std.lv = TRUE, estimator = 'MLMV', data = arm1)

## Reliability

Calphamysubscale1 <- psych::alpha(arm1[,2:5])
omegamysubscale1 <- ci.reliability(data=arm1[,2:5], type="omega", conf.level = 0.95,
               interval.type="perc", B=5)

Calphamysubscale2 <- psych::alpha(arm1[,6:9])
omegamysubscale2 <- ci.reliability(data=arm1[,6:9], type="omega", conf.level = 0.95,
               interval.type="perc", B=5)
```

```{r mod-conv-validity}
modconv <- '

AE =~ spaq_1 + spaq_2 + spaq_3 + spaq_4 
WILL =~ spaq_5 + spaq_6 + spaq_7 + spaq_8 

isi =~ isi_1a + isi_1b + isi_1c + isi_2 + isi_3 + isi_4 + isi_5

dep =~ hads_2 + hads_4 + hads_6 + hads_8 + hads_10 + hads_12 + hads_14

anx =~ hads_1 + hads_3 + hads_5 + hads_7 + hads_9 + hads_11 + hads_13

'

fitconv <- lavaan::sem(model = modconv,
                        std.lv = TRUE, estimator = 'MLMV', data = convdata)

cormat <- cov2cor(lavInspect(fitconv, "cov.lv"))
```

```{r longitudinal invariance, include=FALSE}

data_longitudinal <- mydata |> 
  dplyr::filter(!record_id %in% c(1651, 2015, 2938, 3793)) |>  # dupes
  dplyr::filter(!dplyr::if_all(dplyr::starts_with("spaq_"), ~ is.na(.))) |> 
  dplyr::mutate(across(spaq_5:spaq_8, ~ 6 - .x)) |> 
  dplyr::select(record_id, paste0("spaq_", 1:8), redcap_event_name) |> 
  tidyr::pivot_longer(cols = spaq_1:spaq_8, names_to = "item", values_to = "score") |> 
  dplyr::mutate(item = ifelse(redcap_event_name == "elegibilidade_arm_1", paste0(item, ".1"), paste0(item, ".2"))) |> 
  dplyr::select(-redcap_event_name) |> 
  tidyr::pivot_wider(names_from = item, values_from = score)


configural_model <- '

AE.1 =~ spaq_1.1 + spaq_2.1 + spaq_3.1 + spaq_4.1 
WILL.1 =~ spaq_5.1 + spaq_6.1 + spaq_7.1 + spaq_8.1 

AE.2 =~ spaq_1.2 + spaq_2.2 + spaq_3.2 + spaq_4.2 
WILL.2 =~ spaq_5.2 + spaq_6.2 + spaq_7.2 + spaq_8.2 

'
longFacNames <- list (AE = c("AE.1", "AE.2"),
                      WILL = c("WILL.1", "WILL.2"))

syntax.config <- measEq.syntax(configural.model = configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames)

mod.config <- as.character(syntax.config)

configural.fit <- cfa(mod.config, 
                      data = data_longitudinal, 
                      estimator = "MLMV", 
                      std.lv = TRUE)

syntax.metric <- measEq.syntax(configural.model = configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames,
                               long.equal  = c("thresholds","loadings"))

mod.metric <- as.character(syntax.metric)

fit.metric <- cfa(mod.metric, data = data_longitudinal, estimator = "MLMV",
                  parameterization = "theta", std.lv = TRUE)

conf.met <- compareFit(configural.fit, fit.metric)
summary(conf.met)

lavTestLRT(configural.fit, fit.metric, method="satorra.bentler.2010")

syntax.scalar <- measEq.syntax(configural.model = configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames,
                               long.equal  = c("thresholds","loadings",
                                               "intercepts"))
mod.scalar <- as.character(syntax.scalar)
fit.scalar <- cfa(mod.scalar, data = data_longitudinal, estimator = "MLMV",
                  parameterization = "theta", std.lv = TRUE)

syntax.strict <- measEq.syntax(configural.model = configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames,
                               long.equal  = c("thresholds","loadings",
                                               "intercepts", "residuals"))
mod.strict <- as.character(syntax.strict)
fit.strict <- cfa(mod.strict, data = data_longitudinal, estimator = "MLMV",
                  parameterization = "theta", std.lv = TRUE)

lavTestLRT(configural.fit, fit.metric, fit.scalar, fit.strict, method = "satorra.bentler.2010")
```

```{r longitudinal-fit-differences}
inv_long <- anova(configural.fit, fit.metric, fit.scalar, fit.strict, 
                  method = "satorra.bentler.2010") |> 
  tibble::as.tibble()

fit_stats_strict <- fitmeasures(fit.strict, 
                                fit.measures = c("rmsea.robust", "cfi.robust", 
                                                 "aic", "bic"))

fit_stats_scalar <- fitmeasures(fit.scalar, 
                                fit.measures = c("rmsea.robust", "cfi.robust", 
                                                 "aic", "bic"))

fit_stats_metric <- fitmeasures(fit.metric, 
                                fit.measures = c("rmsea.robust", "cfi.robust", 
                                                 "aic", "bic"))

fit_stats_configural <- fitmeasures(configural.fit, 
                                fit.measures = c("rmsea.robust", "cfi.robust", 
                                                 "aic", "bic"))

long_fit_diffs <- tibble::enframe(c(
        fit_stats_configural - fit_stats_metric,
        fit_stats_metric - fit_stats_scalar,
        fit_stats_scalar - fit_stats_strict)) |> 
  dplyr::mutate(model = rep(c("metric", "scalar", "strict"), each=4),
                .before = name) |> 
  tidyr::pivot_wider(names_from = "name", values_from = "value")
```

```{r group invariance, eval=FALSE, include=FALSE}

syntax.config.group <- measEq.syntax(configural.model = mod.spaq,
                               data = arm1[,2:9],
                               group = "group",
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016")
mod.config.group <- as.character(syntax.config.group)

configural.fit.group <- cfa(mod.config.group, 
                           group = "group",
                           data = dbas[,2:18],
                           estimator = "MLMV",
                           std.lv = TRUE)

```

# Introduction

Acceptance is essential as a therapeutic process for dealing with sleep problems. The Sleep Problem Acceptance Questionnaire (SPAQ) is an 8-itens psychometric instrument that measures the acceptance of sleep difficulties. The SPAQ contains the subscales "Activity Engagement" and "Willingness," with four items composing each factor; respondents rate them on a 7-point scale, where 0 means “Disagree,” and six indicates "Completely
agree." 

## Objective

This study developed a Brazilian-Portuguese of the Sleep Problem Acceptance Questionnaire, examined its factor structure, reliability, and construct validity

# Methods

Data were collected from online surveys responded by 1352 participants, aged 18 to 59 years, with and without insomnia symptoms. Participants completed self-report questionnaires, including:

1. Hospital Anxiety and Depression Scale (HADS);
2. Insomnia Severity Index (ISI);
3. Dysfunctional Beliefs and Attitudes about Sleep Scale (DBAS-16);
4. Acceptance and Action Questionnaire (AAQ-II). 

The SPAQ was translated into Portuguese by three independent translators. The translations were synthesized into a single version by a committee composed of insomnia experts. A pilot study with 15 participants demonstrated that the target population understood the content of the scale.

We ran a Confirmatory Factor Analysis to obtain fit statistics and parameter estimates, using the Diagonally Weighted Least Squares (DWLS) estimator, considering items to be ordered. 

Reliability indices for each of the facets were obtained using Cronbach’s $\alpha$ and McDonald’s $\omega$. We assessed construct validity by examining the latent correlations with variables known to be linked with insomnia.

We also examined whether our scale’s psychometric properties were equal across time points, comparing baseline test scores to a second administration taken 14 days later.

# Results

The 2-factor model showed a good fit [$\chi^2$(`r fitMeasures(fit.spaq, "df")`) = `r round(fitMeasures(fit.spaq, "chisq"),2)`, RMSEA = `r round(fitMeasures(fit.spaq, "rmsea"),3)` [`r round(fitMeasures(fit.spaq, "rmsea.ci.lower"),2)`, `r round(fitMeasures(fit.spaq, "rmsea.ci.upper"),3)`]; CFI = `r round(fitMeasures(fit.spaq, "cfi"),3)`; RNI = `r round(fitMeasures(fit.spaq, "rni"),3)`; TLI = `r round(fitMeasures(fit.spaq, "tli"),3)`]. 

Internal consistency was good for Activity Engagement ($\alpha$ = `r pull(round(Calphamysubscale1$total[1],2))` [`r pull(round(Calphamysubscale1$feldt[[1]],2))`, `r pull(round(Calphamysubscale1$feldt[[3]],2))`], $\omega$ = `r round(omegamysubscale1$est,2)` [`r round(omegamysubscale1$ci.lower,2)`, `r round(omegamysubscale1$ci.upper,2)`]) and Willingness ($\alpha$ = `r pull(round(Calphamysubscale2$total[1],2))` [`r pull(round(Calphamysubscale2$feldt[[1]],2))`, `r pull(round(Calphamysubscale2$feldt[[3]],2))`], $\omega$ = `r round(omegamysubscale2$est,2)` [`r round(omegamysubscale2$ci.lower,2)`, `r round(omegamysubscale2$ci.upper,2)`]). 

The two subscales were weakly and negatively correlated (*r* = `r round(cormat[2,1], 3)`). Both factors also showed negative moderate to strong correlations with measures of insomnia severity, anxiety and depression.




```{r table1, eval=FALSE, include=FALSE}
table1 %>% 
  knitr::kable(
    caption = "Study participants (N = 1419) descriptive statistics.", 
    #align = 'c',
    col.names = c("", "N/Mean (%/SD)"),
    align = "lc",
    table.attr = "style='width:100%;'"
  ) %>% 
  column_spec(1, width = "60%") %>%
  column_spec(2, width = "100%") %>%
  #row_spec(0, background = "#CCCCCC") %>%
  #row_spec(1:nrow(table1), background = "FFFFFF") %>% 
  kableExtra::pack_rows("Gender", 2,3) %>% 
  kableExtra::pack_rows("Race/ethnicity", 4,7) %>% 
  kableExtra::pack_rows("Group", 8,9) %>% 
  kableExtra::pack_rows("Education", 10,12) %>% 
  kableExtra::pack_rows("Region of origin", 13,18) %>% 
  kable_minimal(html_font = "Rasa")
```

<br>

```{r}
covmat <- cov2cor(lavInspect(fitconv, "cov.lv"))
covmat[upper.tri(covmat, diag = FALSE)] <- NA
covmat_t <- corrr::fashion(covmat, decimals = 3)

covmat_t2 <- covmat_t[,-c(5:7)]

rownames(covmat_t) <- c("1. AE", "2. WILL", "3. ISI", "4. Depression", "5. Anxiety")
colnames(covmat_t) <- c("1", "2", "3", "4", "5")

cor_df <- tibble::as_tibble(covmat_t, rownames="Variable")

cor_df |> 
  `colnames<-`(c("Variable", "1", "2", "3", "4", "5")) |> 
  kbl(#booktabs = TRUE, escape = FALSE,
      caption = "Latent correlations and internal consistency levels.", 
      table.attr = "style='width:100%;'",
      align = "llllll")
```

<br>

```{r fit-differences-invariance}


fitdiffs_df <- inv_long[, 5:7] |> 
  dplyr::mutate(
    model = c("configural", "metric", "scalar", "strict"),
    .before = dplyr::everything()) |> 
  dplyr::left_join(long_fit_diffs, by = "model") 

fitdiffs_df[1, -1] <- t(as.data.frame(
  fitmeasures(
    configural.fit, fit.measures = c(
      "chisq.scaled", "df.scaled", "pvalue.scaled",
      "rmsea.robust", "cfi.robust", "aic", "bic"))
))


fitdiffs_df |> 
  dplyr::mutate(
    `Pr(>Chisq)` = ifelse(`Pr(>Chisq)` < .001, "<.001", 
                          weights::rd(`Pr(>Chisq)`, 2)),
    dplyr::across(c(aic, bic), ~round(.x, 0)),
    dplyr::across(c(rmsea.robust, cfi.robust),
                  ~weights::rd(.x, 4)),
    `Chisq diff` = round(`Chisq diff`, 2),
    model = stringr::str_to_title(model)
    ) |> 
  `colnames<-`(c("Model", "$\\Delta\\chi^2$", 
                 "$\\Delta$df", "$\\it{p}$-value", 
                 "$\\Delta$RMSEA", "$\\Delta$CFI",
                 "$\\Delta$AIC", "$\\Delta$BIC")) |> 
  kbl(#booktabs = TRUE, escape = FALSE,
      caption = " Model fit differences for CFA longitudinal measurement invariance tests."
      ) |> 
  footnote(general = "Differences to each preceding constraint level.",
           footnote_as_chunk = TRUE,
           general_title = "Note. ")


```

```{r graph, fig.width=8, fig.height=12, dpi=300}
cfa_plot <- semPaths(fit.spaq, residuals=FALSE,sizeMan=6, rotation =2, groups = "latents", 
         color=wesanderson::wes_palette("GrandBudapest2", n=2), layout = "tree",
         posCol=c("skyblue4"),"std", edge.label.cex = 1,  curvePivot = TRUE,
         nodeLabels = c("S1", "S2", "S3", "S4",
                        "S5", "S6", "S7", "S8",
                        "AE", "WIL")
         )
```




# Conclusion

These findings suggest that the SPAQ is a valid tool for assessing the acceptance of sleep with a Brazilian-Portuguese speaking population. However, due to the low correlation between the two subscales, researchers should refrain from interpreting a total sum score and consider these as individual factors.

## Items

  1. I try to control my sleep by forcing it to happen or being too rigid about my sleep schedule.
	 
  2. I avoid making plans or doing things I enjoy or need to do because of my sleep.
	 
  3. The thought of going to sleep makes me uneasy.
	 
  4. I have negative feelings towards my sleep.
	 
  5. My worries, thoughts, and feelings about my sleep interfere with how I go about my daily life.
	 
  6. I am scared of not being able to fall asleep.
