---
main_topsize: 0.13 #percent coverage of the poster
main_bottomsize: 0.07
poster_height: "150cm"
poster_width: "80cm"
body_textsize: "46px"
#ESSENTIALS
title: '**Development of the Sleep Acceptance Scale (SAS)**'
author:
  - name: '**Marwin M. I. B. Carmo**'
    affil: 1
    main: true
    #orcid: '0000-0002-1099-3857'
    twitter: marwincarmo 
    email: marwin@usp.br
  - name: Maria Laura Nogueira Pires
    affil: 2
    main: true
    email: 'laurapires@actigrafia.org'
  - name: Leo Paulos-Guarnieri
    affil: 3
    main: true
    email: 'guarnieri.leo@gmail.com'
  - name: Renatha El Rafihi Ferreira
    affil: 1
    main: true
    email: 'rerafihi@usp.br'
affiliation:
  - num: 1
    address: Department of Psychiatry, University of São Paulo, Brazil
  - num: 2
    address: Private practice in sleep psychology
  - num: 3
    address: Department of Psychology, Pontifícia Universidade Católica de São Paulo, Brazil
# Main Colour Scheme
primary_colour: "#024F84"
secondary_colour: "#C00000"
body_bgcol: "#ffffff"
#---POSTER BODY OPTIONS---#
reference_textsize: "32px"
caption_textsize: "32px"
affiliation_textsize: "32px"
#Column
column_numbers: 2
#Main Section
main_fontfamily: "PT Sans"
main_textalign: "left"
main_textsize: "160px"
main_textalpha: "cf"
main_findings:
  - "Preliminary evidence for a new measure \nof **Sleep Acceptance**: a promising tool for clinicians and researchers."
main_textcol: "#ffffff"
logoleft_name: img/fapesp-white.png
logoright_name: img/ipq-white3.png
logocenter_name: img/qrcodeSAS.svg
output: 
  posterdown::posterdown_betterport:
    highlight: NULL
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
    template: custom_template.html
bibliography: references.bib
link-citations: true
csl: apa7.csl
#nocite: '@*'
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
#options(knitr.table.format = "html") 
#options(kableExtra.html.bsTable = T)
```

```{r data, include=FALSE}
library(dplyr)
library(ggplot2)
library(modelsummary)
library(huxtable)
library(kableExtra)
library(psych)
library(lavaan)
library(MBESS)
library(tidyr)
library("stringr")
library("DiagrammeR")
library("semPlot")
library("corrr")
```

```{r dataset, include=FALSE}
ebas_full <- readr::read_csv("ebas.csv") %>% 
  dplyr::mutate(ebas_score = dplyr::select(., ebas_1:ebas_6) %>% rowSums())
ebas <- dplyr::select(ebas_full, ebas_1:ebas_6)
table1 <- readr::read_csv("table1.csv")
```

```{r efa, eval=FALSE, include=FALSE}
ebas.pa <- fa.parallel(ebas , cor="poly")
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# CFA ####

# specify the model
CFA.ebas2 <- '
# factor structure
Avoidance =~ ebas_1 + ebas_2
Distress =~ ebas_3 + ebas_4 + ebas_5 + ebas_6
'


# fit the model

# to use an estimator better suited to ordinal or binary items (WLSMV), use "ordered" specification as here: http://lavaan.ugent.be/tutorial/cat.html)
# alternatively, consider using estimator = "MLR" (robust maximum likelihood) for ordinal items with e.g. 7-point response scales


#fitCFA.ebas <- lavaan::cfa(CFA.ebas1, data=ebas, std.lv = TRUE, estimator = 'MLR')
fitCFA.ebas2 <- lavaan::cfa(CFA.ebas2, data=ebas, std.lv = TRUE, estimator = 'DWLS', ordered = TRUE)
# model summary
#summary(fitCFA.ebas, standardized=TRUE, fit.measures = TRUE)
# coefficients only
# coef(fitCFA.mydata)
# CFA diagram from psych package
# lavaan.diagram(fitCFA.mydata, errors=TRUE)
#delta.ebas <- anova(fitCFA.ebas, fitCFA.ebas2)

## Reliability

Calphamysubscale1 <- psych::alpha(ebas[,1:2]) 
omegamysubscale1 <- ci.reliability(data=ebas[,1:2], type="omega", conf.level = 0.95,
               interval.type="perc", B=100)

Calphamysubscale2 <- psych::alpha(ebas[,3:6]) 
omegamysubscale2 <- ci.reliability(data=ebas[,3:6], type="omega", conf.level = 0.95,
               interval.type="perc", B=100)
```

```{r, include=FALSE}
sas.igi <- cor.test(ebas_full$ebas_score, ebas_full$igi_escore)
sas.dbas <- cor.test(ebas_full$ebas_score, ebas_full$dbas_score )
sas.aaq <- cor.test(ebas_full$ebas_score, ebas_full$aaq_score )
sas.hadsa <- cor.test(ebas_full$ebas_score, ebas_full$ehad_ansiedade_escore )
sas.hadsd <- cor.test(ebas_full$ebas_score, ebas_full$ehad_depressao_escore )
sas.spaq <- cor.test(ebas_full$ebas_score, ebas_full$spaq_score )
```

# Introduction

Insomnia symptoms are associated with higher levels of <b><span style="color: #C00000;">metacognitive control</b></span>. Controlled information processing interferes with a healthy sleep process by preventing cognitive deactivation, which involves diminished verbal regulation and control, and acceptance of spontaneous physiological and mental processes. Because general tools for measuring acceptance do not address sleep-related behaviors, the development of specific instruments is essential for studying acceptance of insomnia. 

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

## Objective

This study aimed to develop a new measure of sleep acceptance, titled Sleep Acceptance Scale (SAS), and examine its factorial structure, internal consistency, and construct validity.

# Methods

The SAS was developed as a 6-item self-report questionnaire rated using a scale ranging from 1 (never) to 7 (always), such that high scores indicate lower sleep acceptance.

Data were collected from online surveys responded by 1419 participants, aged 18 to 59 years, with and without insomnia symptoms. Participants completed self-report questionnaires, including:

1. Hospital Anxiety and Depression Scale (HADS);
2. Insomnia Severity Index (ISI);
3. Dysfunctional Beliefs and Attitudes about Sleep Scale (DBAS-16);
4. Acceptance and Action Questionnaire (AAQ-II). 

We conducted a parallel analysis and exploratory factor analysis (with oblimin rotation) to estimate the number of facets underlying the instrument items. 

We ran a Confirmatory Factor Analysis to obtain fit statistics and parameter estimates, using the Diagonally Weighted Least Squares (DWLS) estimator, considering items to be ordered. 

Reliability indices for each of the facets were obtained using Cronbach’s $\alpha$ and McDonald’s $\omega$. We assessed construct validity by examining the relations with variables known to be linked with insomnia.

# Results

The parallel analysis procedure indicated two latent factors: <b><span style="color: #C00000;">Avoidance</b></span> (two items) and <b><span style="color: #C00000;">Distress</b></span> (four items). The 2-factor model showed a good fit [$\chi^2$(`r fitMeasures(fitCFA.ebas2, "df")`) = `r round(fitMeasures(fitCFA.ebas2, "chisq"),2)`, RMSEA = `r round(fitMeasures(fitCFA.ebas2, "rmsea"),3)` [`r round(fitMeasures(fitCFA.ebas2, "rmsea.ci.lower"),2)`, `r round(fitMeasures(fitCFA.ebas2, "rmsea.ci.upper"),3)`]; CFI = `r round(fitMeasures(fitCFA.ebas2, "cfi"),3)`; RNI = `r round(fitMeasures(fitCFA.ebas2, "rni"),3)`; TLI = `r round(fitMeasures(fitCFA.ebas2, "tli"),3)`]. 

Internal consistency was suboptimal for Avoidance ($\alpha$ = `r pull(round(Calphamysubscale1$total[1],2))` [`r pull(round(Calphamysubscale1$feldt[[1]],2))`, `r pull(round(Calphamysubscale1$feldt[[3]],2))`], $\omega$ = `r round(omegamysubscale1$est,2)` [`r round(omegamysubscale1$ci.lower,2)`, `r round(omegamysubscale1$ci.upper,2)`]) and excellent for Distress ($\alpha$ = `r pull(round(Calphamysubscale2$total[1],2))` [`r pull(round(Calphamysubscale2$feldt[[1]],2))`, `r pull(round(Calphamysubscale2$feldt[[3]],2))`], $\omega$ = `r round(omegamysubscale2$est,2)` [`r round(omegamysubscale2$ci.lower,2)`, `r round(omegamysubscale2$ci.upper,2)`]). 

The SAS was positively correlated with ISI (*r* = `r round(sas.igi$estimate,2)`), DBAS-16 (*r* = `r round(sas.dbas$estimate,2)`), AAQ-II (*r* = `r round(sas.aaq$estimate,2)`), HADS-A (*r* = `r round(sas.hadsa$estimate,2)`), HADS-D (*r* = `r round(sas.hadsd$estimate,2)`), and negatively correlated with SPAQ (*r* = `r round(sas.spaq$estimate,2)`).


```{r table1}
table1 %>% 
  knitr::kable(
    caption = "Study participants (N = 1419) descriptive statistics.", 
    #align = 'c',
    col.names = c("", "N/Mean (%/SD)"),
    align = "lc",
    table.attr = "style='width:100%;'"
  ) %>% 
  column_spec(1, width = "60%") %>%
  column_spec(2, width = "100%") %>%
  #row_spec(0, background = "#CCCCCC") %>%
  #row_spec(1:nrow(table1), background = "FFFFFF") %>% 
  kableExtra::pack_rows("Gender", 2,3) %>% 
  kableExtra::pack_rows("Race/ethnicity", 4,7) %>% 
  kableExtra::pack_rows("Group", 8,9) %>% 
  kableExtra::pack_rows("Education", 10,12) %>% 
  kableExtra::pack_rows("Region of origin", 13,18) %>% 
  kable_minimal(html_font = "Rasa")
```

<br>

```{r}
sas_cor <- ebas %>%
  correlate(method = "pearson") %>%
  shave() %>%
  dplyr::mutate(dplyr::across(ebas_1:ebas_6, ~round(.x,2) |> as.character())) |> 
  dplyr::mutate(dplyr::across(everything(), ~replace_na(.x, ""))) |> 
  dplyr::select(-ebas_6) |> 
  dplyr::mutate(term=stringr::str_remove_all(term, "[^0-9]"))

sas_ms <- ebas |> 
  tidyr::pivot_longer(cols = everything(), names_to = "item", values_to = "score") |> 
  dplyr::with_groups(item, summarise, Mean = mean(score), SD = sd(score)) |> 
  dplyr::mutate(dplyr::across(Mean:SD, ~round(.x,2) |> as.character())) |> 
  select(-item)

bind_cols(sas_cor, sas_ms) |> 
  kable(col.names = c("Item", "1", "2", "3", "4", "5", "Mean", "SD"),
        table.attr = "style='width:100%;'",
        align = "c",
        caption="Descriptive statistics and item inter-correlations of the SAS.")
```



```{r include=FALSE}
# To describe edges, we need the origin, path-type, endpoint, and label.
paths <- fitCFA.ebas2 %>%
  parameterestimates %>%
  select(lhs, op, rhs, est)

# Latent variables are left-hand side of "=~" lines
latent <- paths %>%
  filter(op == "=~") %>%
  select(id = lhs) %>%
  distinct %>%
  mutate(shape = "circle")

# Manifest variables are not latent variables
`%not_in%` <- Negate(`%in%`)
manifest <- paths %>%
  filter(op != "~1", lhs %not_in% latent$nodes) %>%
  select(id = lhs) %>%
  distinct %>%
  mutate(shape = "square")

# Nodes are prepared
node_set <- bind_rows(latent, manifest)

# Edges will be labeled by the parameter estimates
all_paths <- paths %>%
  filter(op != "~1") %>%
  mutate(label = round(est, 2)) %>%
  select(-est)

# Factor loadings are the paths in the "=~" lines
loadings <- all_paths %>%
  filter(op == "=~") %>%
  mutate(from = lhs, to = rhs, style = "solid") %>%
  select(from, to, style, label)

# This is now an edge dataframe
loadings
```


```{r include=FALSE}
# Regressions are the paths in the "~" lines
regressions <- all_paths %>%
  filter(op == "~~") %>%
  rename(to = lhs, from = rhs) %>%
  mutate(style = "dashed") %>%
  select(from, to, style, label)

edge_set <- bind_rows(loadings, regressions)
edge_set
```





```{r out.width="80%", dpi=300, fig.cap="Structural equation model for the Confirmatory Factor Analysis for the final version of the Sleep Acceptance Scale. Avd = Avoidance, Dis = Distress."}
grViz("
digraph {

  rankdir=LR;
  
  'Avd' [shape = 'circle' height=1.2 fontsize=30]
  'Dis' [shape = 'circle' height=1.2 fontsize=30]
  'x1' [shape = 'square']
  'x2' [shape = 'square']
  'x3' [shape = 'square']
  'x4' [shape = 'square']
  'x5' [shape = 'square']
  'x6' [shape = 'square']

  'e1' [shape = 'circle']
  'e2' [shape = 'circle']
  'e3' [shape = 'circle']
  'e4' [shape = 'circle']
  'e5' [shape = 'circle']
  'e6' [shape = 'circle']

  'Avd'->'x1' [ label = '.55']
  'Avd'->'x2' [ label = '.72']
  'Dis'->'x3' [ label = '.93']
  'Dis'->'x4' [ label = '.95']
  'Dis'->'x5' [ label = '.85']
  'Dis'->'x6' [ label = '.89']
  
  'x1' -> 'e1' [dir=back rank=same]
  'x2' -> 'e2' [dir=back]
  'x3' -> 'e3' [dir=back]
  'x4' -> 'e4' [dir=back]
  'x5' -> 'e5' [dir=back]
  'x6' -> 'e6' [dir=back]

  'e1':e -> 'e1':e [dir=both label = '.70']
  'e2':e -> 'e2':e [dir=both label = '.48']
  'e3':e -> 'e3':e [dir=both label = '.13']
  'e4':e -> 'e4':e [dir=both label = '.10']
  'e5':e -> 'e5':e [dir=both label = '.28']
  'e6':e -> 'e6':e [dir=both label = '.21']

  'Avd':w -> 'Dis':w [dir=both label = .79 constraint=false]

}
")
```


# Conclusion

Preliminary evidence suggest that the SAS might be a reliable and valid indicator of acceptance of sleep problems in people with and without sleep problems.

## Items

1. Tento controlar meu sono forçando que ele aconteça, ou sendo muito rígido(a) em relação aos horários de dormir e acordar.

2.	Evito compromissos ou deixo de fazer coisas que gosto ou preciso por causa do meu sono.

3.	Sinto-me desconfortável ao pensar no momento de dormir.

4.	Tenho sentimentos negativos em relação ao meu sono.

5.	Minhas preocupações, pensamentos e sentimentos sobre o meu sono interferem na maneira como eu lido com o meu dia a dia.

6.	Tenho medo de não dormir.
