
# Method

## Participants and Study Design

```{r power-analysis}
ss_closefit_dbas <- semTools::findRMSEAsamplesize(rmsea0=.05, rmseaA=.059, df=98, power=.80)
ss_notclosefit_dbas <- semTools::findRMSEAsamplesize(rmsea0 = .08, rmseaA = .059, df=98, power=.80)

ss_closefit_spaq <- semTools::findRMSEAsamplesize(rmsea0=.05, rmseaA=.08, df=19, power=.80)
ss_notclosefit_spaq <- semTools::findRMSEAsamplesize(rmsea0 = .10, rmseaA = .08, df=19, power=.80)

```

To estimate an adequate sample size for the confirmatory factor analyses (CFAs) we used MacCallum et al.’s [-@maccallum1996] root-mean-square error of approximation (RMSEA) tests of close and not-close fit. All tests were conducted in R 4.1.3 [@R-base] using `semTools` version `r packageVersion("semTools")` [@semtools]. Morin [-@morin2007a] reports a RMSEA of 0.059 in a CFA for DBAS-16. Taking this value as prior guess for the true RMSEA score, we calculated the sample sizes required to to reject the test for not-close fit of RMSEA > 0.08 and the test of close fit of RMSEA < 0.05 with a power of 0.80 and $\alpha$ of 0.05. Results show that `r ss_notclosefit_dbas` subjects are necessary to reject the test for not-close fit, and `r ss_closefit_dbas` participants are required for rejection of the test of close fit. Therefore, we aimed at a minimum sample size of `r ss_closefit_dbas` participants. SPAQ's fit index was not considered in this power analysis due to the large RMSEA (0.081) reported originally [@bothelius2015].

This study was approved by the Ethics Committee of the General Hospital of the University of São Paulo, School of Medicine (HC-FMUSP), São Paulo, Brazil (CAAE: 46284821.1.0000.0068). Inclusion criteria was age between 18 and 59 years and reporting no difficulties in reading or writing in Portuguese. 

Participants were recruited mainly from advertisement on the internet, especially on HC-FMUSP's social media platforms (Instagram and Facebook). The data collection took place between May 2021 through July 2022, with brief breaks in between. Because the measures evaluated in this study refer to sleep difficulties, we sought to include participants both with and without insomnia complaints. The first group was composed by people registered for an experimental behavioral treatment for insomnia, which was also organized by the Department of Psychiatry of HC-FMUSP and which this study is a branch of. To recruit participants without insomnia complaints we asked for volunteer participation from people believing not having sleeping problems.

Bad sleepers were classified according to the presence of insomnia complaints: (i) difficulty initiating and/or maintaining sleep, defined as a sleep onset latency and/or wake after sleep onset greater than or equal to 30 minutes, with a corresponding sleep time of less than or equal to six hours per night; (ii) presence of insomnia for more than three nights per week and more than three months; (iii) sleep disturbance (or associated daytime fatigue) causing significant distress or impairment in social, occupational, or other areas of functioning. This definition represents a combination of criteria from the American Academy of Sleep Medicine, the International Classification of Sleep Disorders, and the Diagnostic and Statistical Manual of Mental Disorders, along with quantitative cutoffs typically used in insomnia research **(American Academy of Sleep Medicine, 2014; American Psychiatric Association, 2013; Edinger, et al., 2004)**. In addition to these criteria, participants total score on the Insomnia Severity Index should not exceed 7 points [@bastien2001].

Participants were informed about the main objective of the research and signed the informed consent. They were informed that their answers would be kept confidential, and that all procedures
guaranteeing the privacy of their results would be adopted. Then, they were requested to respond to an online survey using REDCap electronic data capture tools [@harris2009research; @harris2019redcap], including the Brazilian-Portuguese versions of DBAS-16 and SPAQ and other auxiliary instruments.

### Item translation

We mainly based our methods on Beaton's ([-@beaton2000]) recommendations with the addition of more up to date insights from @borsaAdaptacaoValidacaoInstrumentos2012. Figure \@ref(fig:flowcca) summarize the steps taken in the process.

The following procedures were applied both to DBAS-16 as well as to SPAQ. Only the expert committee and the first translation team had a different configuration for each instrument.

In the first stage the items of the original versions were translated from English (source language) to Portuguese (target language) by three independent translators, of which two were familiar with the instrument constructs and the other English teachers unaware of the instrument concepts and with no clinical or medical background. The three versions were synthesized by an expert committee of health professionals experts in insomnia. A form adapted from @koller2012 was given to each member of the committee to register the rationale for the decisions (see Appendix \ref{appendix:form}). Then, two independent translators native speakers of the source language back translated the synthesized version to English. We reconciled the back translations into a single version and submitted it to appreciation by both first authors of the original questionnaires. Together with the expert committee we debated over suggestions raised by the original authors and made changes accordingly to the translated version. 

At the final step, we conducted a pilot study with 15 participants from the target population to probe the pre-final version. There were 12 female participants and overall mean age was 43 years (range: 19--57 years). To prevent restricting feedback to specific regional contexts [@borsaAdaptacaoValidacaoInstrumentos2012], we recruited participants from the five Brazilian regions and with varying educational levels. We were able to interview nine participants from the Southwest region, three from South, two from Northeast and one from Middle-west. We conducted individual cognitive interviews with each participant.

```{r flowcca, echo = FALSE, message=FALSE, fig.align='center', fig.cap='Stages of cross-cultural adaptation. Adapted from “Cross-Cultural Adaptation and Validation of Psychological Instruments: Some Considerations”, by J. C. Borsa, B. F. Damásio and D. R. Bandeira, 2012, Paidéia, 22(53), 423-432; “Guidelines for the Process of Cross-Cultural Adaptation of Self-Report Measures”, by D. E. Beaton, C. Bombardier, F. Guillemin and M. B. Ferraz, 2000, SPINE, 25(24), 3186-3191.'}
knitr::include_graphics("cca_diagram.pdf")
```

## Aditional measures

1. *Insomnia Severity Index (ISI)* [@bastien2001; @morin2011a] is a 7-item questionnaire to assess insomnia severity and its impact on the patient's life. Raters use a 5-point scale ranging from 0 (no problem) to 4 (very severe problem). We used the Brazilian-Portuguese version [@castro].

2. *The Hospital Anxiety and Depression Scale (HADS)* [@zigmond1983hospital] is a scale used to assess psychological distress in non-psychiatric patients. It is formed by a two-factor structure with seven items assessing Anxiety plus seven other items measuring Depression. A Brazilian-Portuguese version produced by @botega1995transtornos was used.

3. *Acceptance and Action Questionnaire-II (AAQ-II)* [@bond2011preliminary; @hayes2004measuring] is a measure of psychological flexibility composed by seven items rated in a scale from 1 (never true) to 7 (always true). it is scored by adding up scores for each question. Higher scoring indicate less flexibility. The Brazilian-portuguese version used in this study was produced by @barbosa2015propriedades.

## Analytical Plan

### Descriptive statistics

This phase comprise examination of response frequency and item statistics in order to assess item variation, distribution and data entry quality. Items with insufficient variation might be bad for differentiating respondents and may need to be excluded or merged into fewer categories [@dima2018]. We'll also diagnose inter-item correlations and scan for multivariate outliers (via Mahalanobis distance) to identify if there are any anomalous response patterns.

### Examination of item properties with IRT

Next, we examine item response patterns using Mokken Scaling Analysis (MSA), a non-parametric Item Response Theory (IRT) thecnique. Explain what MSA is about. 

### Scale structure via factor analysis

### Reliability estimates

### Calculation of global scores
