
# Partial results

```{r packages, include=FALSE}
library(dplyr)
library(kableExtra)
```

```{r dataset, include=FALSE}
raw.data <- readr::read_csv("~/Mestrado/data/TerapiaDeAceitaoECom_DATA_2022-08-17_1702.csv")
gc <- readr::read_csv("~/Mestrado/data/gc1708.csv")
renda_corrigida <- readr::read_csv("G:/Documentos/ProjetosR/ASONO/act_insonia/dados/renda_corrigida_2610.csv")

location <- readr::read_csv("data_analyses/location_db.csv") |> 
  dplyr::select(record_id, estado, region)

dat <- raw.data |>
  # fill idade
  dplyr::group_by(record_id) |>  
  tidyr::fill(nome, .direction = "down") |>
  tidyr::fill(idade, .direction = "down") |> 
  tidyr::fill(cpf, .direction = "down") |> 
  dplyr::ungroup() |> 
  dplyr::mutate(identificao_timestamp = lubridate::ymd_hms(identificao_timestamp),
                idade2 = lubridate::time_length(difftime(identificao_timestamp, data_nascimento, unit="weeks"), "years"),
                age = dplyr::coalesce(idade, idade2)) |> 
  dplyr::select(-c(idade, idade2)) |> 
  dplyr::filter(!is.na(escala_breve_de_aceitao_do_sono_complete),
                age >= 18 & age < 60,
                dbas_16_versao_brasileira_complete == 2) |> 
  dplyr::select(record_id, nome, cpf, redcap_event_name, sexo, age, etnia, estado_civil, escolaridade, renda, ocupacao, igi_escore, ehad_depressao_escore,
                ehad_ansiedade_escore, aaq_score, dbas_score, contains("spaq_"), ebas_1:ebas_6) |> 
  dplyr::filter(redcap_event_name == "elegibilidade_arm_1") |> 
  dplyr::filter(!stringr::str_detect(nome, stringr::regex("teste", ignore_case = TRUE))) |> 
  dplyr::mutate(grupo = dplyr::case_when(cpf %in% gc$cpf ~ 0,
                                         TRUE ~ 1)) |> 
  dplyr::left_join(renda_corrigida, by = "record_id") |> 
  dplyr::mutate(renda.x = case_when(
    !is.na(renda.y) ~ renda.y,
    # renda.x >= 1000000 ~ NA_real_,
    # renda.x == 0 ~ NA_real_,
    TRUE ~ renda.x
    )
         ) |> 
  dplyr::rename(renda = renda.x) |> 
  dplyr::select(-renda.y) |> 
  dplyr::left_join(location, by = "record_id") |> 
  dplyr::mutate(region = dplyr::case_when(
    estado == "Distrito Federal/Goiás" ~ "Região Centro-Oeste",
    TRUE ~ region
  ))
```


## Cross-cultural adaptation

The initial translation of SPAQ and DBAS-16 instructions, rating scale, and items was a mix of translations produced by the three (for each instrument) forward translators. To some items a determined translation was taken with minor or no modifications. Others were a merge of two or more versions with additions were it deemed necessary. The instruments versions produced in each stage of the cross-cultural adaptation process, as well as a detailed documentation of criteria for decisions, are available at [https://osf.io/av45j/](https://osf.io/av45j/).

Once each stage of the translation process was completed, both instruments were submitted to appreciation by a sample of 15 subjects of the target population. Overall, participants had a good comprehension of the test items and instructions and only a single term of the DBAS-16 required alteration for a more natural reading in the target language. The final version of both SPAQ and DBAS-16 are on Appendix X. Colocar mudança nas instruções do SPAQ.

## Sample description

After excluding individuals who did not meet the inclusion criteria and those who failed to complete at least the first questionnaire on the survey (DBAS-16), the final sample was comprised of `r nrow(dat)` individuals, of which `r nrow(dat[dat$sexo == 1,])` were female and `r nrow(dat[dat$grupo == 1,])` reported insomnia symptoms. Sample mean age was `r round(mean(dat$age, na.rm=TRUE),2)` years (SD = `r round(sd(dat$age, na.rm=TRUE),2)`, range: `r round(min(dat$age, na.rm=TRUE),1)`--`r round(max(dat$age, na.rm=TRUE),1)` years). There were `r nrow(dat[dat$ocupacao == 5,])` participants who reported having a formal job, and `r nrow(dat[dat$escolaridade == 9,])` had a university degree. A detailed description of the sample is found on Table X.

```{r include=FALSE}
tab1 <- dat |> 
  dplyr::mutate(renda = dplyr::case_when(
    renda > 40000 | renda < 100 ~ NA_real_,
    TRUE ~ renda
  )) |> 
  dplyr::mutate(escolaridade = dplyr::case_when(
    escolaridade %in% c(1:6) ~ "Primary School",
    escolaridade %in% c(7:8) ~ "Secondary School",
    escolaridade == 9 ~ "University degree or higher"),
    etnia = dplyr::case_when(
    etnia == 1 ~ "White",
    etnia %in% c(2,3) ~ "Black",
    etnia == 4 ~ "Asian",
    !etnia %in% c(1:4) ~ "Other/Not informed"),
    region = dplyr::case_when(
    region == "Região Centro-Oeste" ~ "Central-West",
    region == "Região Nordeste" ~ "Northeast",
    region == "Região Norte" ~ "Northern",
    region == "Região Sudeste" ~ "Southeast",
    region == "Região Sul" ~ "Southern"),
    sexo = dplyr::case_when(
    sexo == 1 ~ "Female",
    sexo == 2 ~ "Male"),
    grupo = dplyr::case_when(
    grupo == 1 ~ "Insomnia",
    grupo == 0 ~ "Good sleeper"),
    ocupacao = dplyr::case_when(
    ocupacao == 1 ~ "Retired",
    ocupacao == 2 ~ "Self-employed",
    ocupacao %in% 3:4 ~ "Unemployed",
    ocupacao == 5 ~ "Regular job", 
    ocupacao == 6 ~ "Informal work",
    ocupacao == 7 ~ "Student"
    )) |> 
  dplyr::select(sexo:ocupacao, grupo, region) %>%
  tableone::CreateTableOne(data=.) 
  #table1::table1(~., data=.)
```


```{r echo=FALSE}
tblheader <- paste0('*n*', " = ", nrow(dat))

tab1b <- print(tab1, printToggle = FALSE) |> 
  as.data.frame() |> 
  slice(-1)
rnames <- rownames(tab1b)
rnames[1] <- "Sex Male (%)"
rnames[2] <- "Age (mean (SD))"
rnames[3] <- "Race (%)"
rnames[8] <- "Marital Status (mean (SD))"
rnames[9] <- "Educational Level (%)"
rnames[13] <- "Monthly income (mean (SD))"
rnames[14] <- "Occupation (%)"
rnames[21] <- "Insomnia (%)"
rnames[22] <- "Region (%)"
row.names(tab1b) <- rnames

kbl(tab1b,
    caption = "Sample description",
    booktabs = TRUE,
    col.names = tblheader, format = 'latex',
    align = "c") |>
  kable_styling(full_width = TRUE) |>
  #column_spec(1, align = "c") |>
  add_indent(c(4:7, 10:12, 15:20, 23:27))
```




