---
title: "Cross-cultural adaptation and psychometric studies of the Dysfunctional Beliefs and Attitudes about Sleep scale and the Sleep Problem Acceptance Questionnaire"
shorttitle:  ""

author: 
  - name        : "Marwin M I B Carmo"
    affiliation : ""

affiliation:
  - id          : ""
    institution : ""

class: doc,11pt, twoside
#lang: english
figsintext: yes
lineno: no
footnotelist: no
figurelist: no
link-citations: true
urlcolor: blue

bibliography: ["Bibs/references.bib", "Bibs/r-references.bib"]
csl: apa7.csl
output:
  papaja::apa6_pdf:
    includes:
      after_body: 
      #- Chapter01_files/Appendix_Chapter1.tex
      - croco/Appendix.tex
      
header-includes:
  - \setcounter{tocdepth}{3}
  #- \addtolength\oddsidemargin {0.5cm}
  #- \addtolength\evensidemargin {-0.5cm}
  - \setlength{\parskip}{5pt}
  - \linespread{1.5} 
  - \usepackage{setspace} 
  - \usepackage{tabu}
  - \usepackage{ragged2e}
  - \usepackage{graphicx}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks = true, urlcolor = blue, linkcolor = blue, citecolor = red}
  - \shorttitle{}
  - \fancyheadoffset[L]{0pt}  
  - \fancyhf{}
  - \fancyhead[RO,LE]{\small\thepage} 
  - \renewcommand{\headrulewidth}{0pt}
  - \interfootnotelinepenalty=10000
---

<!--To remove the first two pages use the following code in ubuntu:
pdftk myDocument.pdf cat 3-end output removedPages.pdf
-->


<!--\clearpage\mbox{}\thispagestyle{empty}\clearpage-->
\newpage
\setcounter{page}{1}

\thispagestyle{empty}
\begin{center}
\vspace*{10mm}
\textbf{\Large Cross-cultural adaptation and psychometric studies of the DBAS-16 and SPAQ}\\
\bigskip
\begin{figure}[h]
\begin{center}
\includegraphics[width=!,totalheight=!,scale=0.18]{usp-brazao.jpg}
\end{center}
\end{figure}
{\setstretch{1.7} 
Qualifying project\\
for\\ 
Obtaining the degree of Master of Science\\
Faculty of Medicine\\
University of São Paulo  \\
Concentration: Psychiatry  \\
presented by  \\ 
\smallskip
\textbf{Marwin M I B Carmo}\\
\smallskip
Thesis advisor:  \\
Dr. Renatha El Rafihi Ferreira  \\
\smallskip
São Paulo, October, 2022\\
}
\end{center}

<!--\clearpage\mbox{}\thispagestyle{empty}\clearpage-->

<!--
\vspace*{\fill}
\thispagestyle{empty}
\begin{flushleft}
1. Berichterstatter: Prof. Dr. XXX (Köln) \\
2. Berichterstatter: Prof. Dr. XXX (Köln) \\  
Diese Dissertation wurde von der Humanwissenschaftlichen Fakultät der Universität zu Köln im Juli 2018 angenommen.\\
Tag der mündlichen Prüfung: XX.XX.2018
\end{flushleft}

\clearpage\mbox{}\thispagestyle{empty}\clearpage
-->
\newpage

\begin{flushleft}
{\setstretch{1.0}
\tableofcontents
}
\end{flushleft}

\newpage

<!--clearpage\mbox{}\thispagestyle{empty}\clearpage-->

\thispagestyle{empty}

# Summary

Insomnia disorder is characterized by frequent complaints about the quality and quantity of sleep. Prolonged exposure may cause physical and psychological damage. Negatively toned activity about sleep is a known reinforcer of insomnia. Because of that, treatments focusing on cognitive components of insomnia, such as Cognitive Behavior Therapy and Acceptance and Commitment Therapy, are popular alternatives or complements to drug therapy with known efficacy. Some specific tools to assess sleep-related cognitions are the Dysfunctional Beliefs and Attitudes about Sleep Scale (DBAS-16) -- measuring the strength of agreement to maladaptive beliefs about sleep--, and the Sleep Problem Acceptance Questionnaire (SPAQ), created to assess acceptance levels of sleep problems. Although these scales have been subject to psychometric scrutiny, it is necessary to test for validity evidence with a Brazilian-Portuguese-speaking sample to achieve valid, reliable, and reproducible results with such a population using these tools. Therefore, the present study proposes a cross-cultural adaptation and study of the psychometric properties and validity evidences of the DBAS-16 and the SPAQ. The target sample is between 18 and 59 years old, with participants with and without insomnia complaints. The steps of the cross-cultural adaptation process were: forward translation, synthesis, back-translation, review and, pre-testing. We collected data from 1397 individuals with a mean age of 38.4 years, of which 1130 were female and 1062 reported insomnia symptoms. The analysis plan includes: Non-parametric item response theory, Confirmatory Factor Analyses, Multiple-group CFA, Reliability estimates of internal consistency and temporal stability, test of convergent validity with related constructs, and exploratory analysis using the Network psychometric approach.

\begin{flushleft}
\emph{Keywords}: sleep-related cognitions, validity, insomnia, assessment.
\end{flushleft}

\newpage

```{r 'Chapter01.Rmd', child="Chapter01.Rmd"}
```

\newpage

```{r 'Chapter03.Rmd', child="Chapter03.Rmd"}
```

\newpage

```{r 'Chapter04.Rmd', child="Chapter04.Rmd"}
```

\newpage

```{r 'Chapter05.Rmd', child="Chapter05.Rmd"}
```

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
