---
title: "The association between sleep-related cognitions, psychological flexibility and insomnia symptoms"
author: "Marwin M. I. B. Carmo"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
bibliography: psicologia_sono.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data, include=FALSE}
dados_c <- readr::read_csv("dadosc.csv")
mod1 <- lm(igi_escore ~ idade + sexo +ehad_ansiedade_escore + ehad_depressao_escore + dbas_score*aaq_score, data = dados_c)
s <- summary(mod1)
```


## Introduction

Psychological disorders like anxiety and depression are known to co-occur with symptoms of insomnia and are possible aggravating factors for sleep problems [@belanger2016]. Excess of negatively toned cognitive activity about sleep may maximize worries, ruminative thoughts can increase arousal and an anxiety state shifts attention to sleep-related threats [@harvey2002].The cognitive-behavioral model of insomnia proposes that arousal, cognitions, behaviors, and perceived consequences play an important role in maintaining insomnia symptoms [@morin1993]. The metacognitive model of insomnia expands this idea by defining cognitive activity as primary arousal and how one interprets those thoughts as a secondary arousal. Primary and secondary processes mutually influence each other in respect that the content of maladaptative sleep-related thoughts can have a greater negative impact in the maintenance of insomnia disorder if one is rigidly attached to them [@ong2012].

### Aim

To investigate if psychological flexibility moderates the relationship between dysfunctional beliefs about sleep and insomnia severity.

## Methods

Data were collected from online surveys responded by 779 adults, age 18 to 59 years, who reported experiencing insomnia symptoms. Participants completed self-report questionnaires, including: Hospital Anxiety and Depression Scale (HADS), Insomnia Severity Scale (ISI), Dysfunctional Beliefs and Attitudes about Sleep Scale (DBAS-16) and Acceptance and Action Questionnaire (AAQ-II). A linear model (estimated using OLS) was fitted to predict ISI scores from DBAS-16 and AAQ-II with age, sex and HADS subscales as covariates.

## Results

All continuous predictors were mean-centered and deviation coding was used for categorical predictors. The results of the regression indicated that the model significantly predicted insomnia severity (F(`r round(s$fstatistic[2], 2)`, `r round(s$fstatistic[3], 2)`) = `r round(s$fstatistic[1], 2)`, p < .001, $R^2_{adj}$ = `r round(s$adj.r.squared, 2)`), accounting for `r round(s$adj.r.squared, 2)*100`% of the variance. 

## Conclusion

## References
