---
main_topsize: 0.01 #percent coverage of the poster
main_bottomsize: 0.07
poster_height: "25.40cm" 
poster_width: "14.29cm"
#ESSENTIALS
title: '**The association between sleep-related cognitions, psychological flexibility, and insomnia symptoms**'
author:
  - name: '**Marwin M. I. B. Carmo**'
    affil: 1
    main: true
    #orcid: '0000-0002-1099-3857'
#    twitter: marwincarmo 
    email: marwin@usp.br
  - name: Renatha El Rafihi Ferreira
    affil: 1
    main: true
    email: 'rerafihi@usp.br '
affiliation:
  - num: 1
    address: Department of Psychiatry, Faculty of Medicine, University of São Paulo, Brazil
# Main Colour Scheme
primary_colour: "#024F84"
secondary_colour: "#C00000"
body_bgcol: "#ffffff"
#---POSTER BODY OPTIONS---#
reference_textsize: "8pt"
caption_textsize: "6pt"
affiliation_textsize: "10pt"
#Column
column_numbers: 2
#Main Section
main_fontfamily: "PT Sans"
main_textalign: "left"
main_textsize: "12pt"
main_textalpha: "cf"
main_findings:
main_textcol: "#ffffff"
logoleft_name: img/fapesp-white.png
logoright_name: img/ipq-white3.png
logocenter_name: img/qrcodeL.svg
output: 
  posterdown::posterdown_betterport:
    highlight: NULL
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
    template: custom_template.html
#bibliography: references_association.bib
link-citations: false
csl: apa7.csl
nocite: '@*'
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
#options(knitr.table.format = "html") 
#options(kableExtra.html.bsTable = T)
```

```{r data, include=FALSE}
library(dplyr)
library(ggplot2)
library(modelsummary)
library(huxtable)
library(kableExtra)
source("plot_template.R")
dados <- readr::read_csv("G:/Documentos/ProjetosR/ASONO/act_insonia/dados/base_arrumada_1409.csv")
table1 <- readr::read_csv("table1data.csv")
```

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

# Introduction

The <b><span style="color: #C00000;">metacognitive model of insomnia</b></span> defines cognitive activity as primary arousal and the interpretation given to those thoughts as a secondary arousal. It proposes that the content of maladaptive sleep-related thoughts contributes to the maintenance of insomnia disorder due to the  <b><span style="color: #C00000;">mutual influence of primary and secondary processes</b></span>. This work tests if psychological inflexibility moderates the relationship between dysfunctional beliefs about sleep and insomnia severity.

# Methods

Data were collected from online surveys responded by 629 adults, aged 18 to 59 years, who reported experiencing insomnia symptoms. Participants completed self-report questionnaires, including: I) Hospital Anxiety and Depression Scale (HADS); II) Insomnia Severity Index (ISI); III) Dysfunctional Beliefs and Attitudes about Sleep Scale (DBAS-16); IV) Acceptance and Action Questionnaire (AAQ-II). A linear model (estimated using OLS) was fitted to predict ISI scores from DBAS-16 and AAQ-II with age, sex and HADS subscales as covariates.

# Results

```{r model, include=FALSE}
dados_c <- dados |> 
  dplyr::mutate(idade = idade - 18) |> 
  dplyr::mutate(dplyr::across(c(igi_escore:spaq_score), ~scale(.x) %>% as.vector)) %>% 
  dplyr::mutate(sexo = ifelse(sexo == "F", .5,-.5),
                sexo = factor(sexo))

mod1 <- lm(igi_escore ~ idade + sexo +ehad_ansiedade_escore + ehad_depressao_escore + dbas_score*aaq_score, data = dados_c)
s <- summary(mod1)
```

The model significantly predicted ISI scores (F(`r round(s$fstatistic[[3]])`, `r round(s$fstatistic[[2]])`) = `r round(s$fstatistic[[1]],2)`), *p* < 0.001, $R^2_{adj}$ = `r round(broom::glance(mod1)$adj.r.squared,2)`), accounting for `r round(broom::glance(mod1)$adj.r.squared,2)*100`% of the variance.

```{r table2, out.width="100%"}

models <- list(
    "Coef."     = mod1,
    "Std.Error" = mod1)
cm <- c( '(Intercept)' = 'Constant', 'idade' = 'Age', 'sexo0.5' = 'Sex',
         'ehad_ansiedade_escore' = 'HADS-A', 'ehad_depressao_escore' = 'HADS-D',
         'dbas_score' = 'DBAS', 'aaq_score' = 'AAQ',
         'dbas_score:aaq_score' = 'DBAS x AAQ')
cap <- 'Table 1: Results from the regression model examining the effects of age, sex, cognitive processes (DBAS-16 and AAQ-II) and anxiety and depressive symptoms on the severity of insomnia.'

tab <- modelsummary(models, output = 'huxtable', align = 'lcr',
                    estimate  = c("{estimate} [{conf.low}, {conf.high}]{stars}", "{std.error}"),
                    coef_map = cm, stars = TRUE, 
                    title = cap, 
                    gof_omit = ".*",
                    statistic = NULL,
                    note = "* p < 0.05, ** p < 0.01, *** p < 0.001",
                    coef_omit = "Intercept") 
tab %>% 
  set_font_size(8) %>% 
  set_bottom_padding(1,2:3, 1) %>% 
  set_col_width(c(0.4,0.5,0.1)) %>%  
  set_align(1, everywhere, "right") %>% 
  #set_align(everywhere, c(2,3), "left") %>%
  set_bold(row = 1, col = everywhere) %>% 
  set_markdown_contents(1, 2, "Coef.") %>% 
  set_markdown_contents(1, 3, "SE") %>%
  set_top_border(row = 1, value = 0) %>% 
  set_bottom_border(row = 9, value = 0) %>% 
  set_background_color(row = c(1,9), col = everywhere, "white") %>% 
  set_width(1)
  #set_background_color(row = 2:nrow(.), col = everywhere, "white") 
  
```

```{r myprettycode, echo=FALSE, fig.cap='Decomposing dysfunctional beliefs about sleep by psychological inflexibility interaction via simple slopes for psychological inflexibility.', fig.height=6, out.width="100%", dpi=300}
#trim whitespace
par(mar=c(2,2,0,0))

dados_c <- dados |> 
  dplyr::mutate(idade = idade - 18) |> 
  dplyr::mutate(dplyr::across(c(igi_escore:spaq_score), ~scale(.x) %>% as.vector)) %>% 
  dplyr::mutate(sexo = ifelse(sexo == "F", .5,-.5),
                sexo = factor(sexo))

mod1 <- lm(igi_escore ~ idade + sexo +ehad_ansiedade_escore + ehad_depressao_escore + dbas_score*aaq_score, data = dados_c)

aaq_high <- dados_c %>% 
  dplyr::mutate(aaq_score = aaq_score + 1)

aaq_low <- dados_c %>% 
  dplyr::mutate(aaq_score = aaq_score - 1)

pred_high <- predict(mod1, aaq_high)
pred_low <- predict(mod1, aaq_low)
pred_avg <-  predict(mod1, dados_c)

figure1 <- dados_c %>% 
  select(igi_escore, dbas_score) %>% 
  mutate(alto = pred_high,
         baixo = pred_low,
         avg = pred_avg) %>% 
  tidyr::pivot_longer(cols = c("alto", "baixo", "avg"), names_to = "grupo", values_to = "igi") %>% 
  ggplot() +
  geom_point(aes(x = dbas_score, y = igi_escore), alpha = .08, position = "jitter") +
  stat_smooth(aes(x = dbas_score, y = igi, color = factor(grupo)), 
              method="lm",se=FALSE, fullrange = TRUE,
              size = 1.3) +
  my_theme +
  scale_color_viridis_d(option = "A", label = c("High (+1SD)", "Average","Low (-1SD)"), 
                        direction = -1, begin = .9, end = .4, name = "AAQ-II") +
  labs(x = "Standardized DBAS-16 scores",
       y = "Standardized ISI scores") +
  theme(legend.position = "bottom",
        axis.title=element_text(size=16))
figure1
```

# Conclusion

The <b><span style="color: #C00000;">significant interaction effect</b></span> indicates that the prediction effect of dysfunctional beliefs about sleep may become more positive for additional levels of psychological inflexibility.
